Randomly-Scoped Lambda Calculus

### 2012-12-20 RSLC/paper/paper.tex
write sect. 4

Unfortunately, I wasn't even able to find an RSLC term that satisfied that identity. I had to also alter the way recursive functions call themselves -- by referring to the combinator itself, passing it the recursive arguments, and {\em then} the substituted version of themselves. Since that makes no sense in words, just have a look:
\begin{eqnarray*}
\Ya &=& \lambda ag. gag \\
\Yab &=& \lambda abg. gabg \\
\Yabc &=& \lambda abcg. gabcg \\
&\dots&
\end{eqnarray*}
Here's a sort of handwavy syntactic transformation example for $\mathsf{fix}$ constructs that supports one, two, etc arguments:
\begin{eqnarray*}
\mathsf{fix}_1(f,a.e) &=& \lambda a. \Ya~a~(\lambda af. [^{(\Ya~x~f)}/_{f(x)}]e) \\
\mathsf{fix}_2(f,a,b.e) &=& \lambda ab. \Ya~a~b~(\lambda abf. [^{(\Ya~x~y~f)}/_{f(x)(y)}]e) \\
&\dots&
\end{eqnarray*}

@@

\newcommand{\addition}{\maltese}
\begin{center}
\begin{tabular}{rcl}
$\addition$ & = &
\hspace{-1em}\multirow{4}{*}{
\begin{tabular}{lllll}
\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
&& $\Zero$ & $\Rightarrow$ & $\lambda f.~n$ \\
&& $\Suc{m_2}$ &$\Rightarrow$& $\lambda f.~\Suc{\Yab~m_2~n~f}$
\end{tabular}
}
\\
&&\\
&&\\
&&\\
&&\\
$\mathsf{add}~M~N$ &=& $\Yab~M~N~\addition$
\end{tabular}
\end{center}

@@

\newcommand{\multiplication}{\bigstar}
\begin{center}
\begin{tabular}{rcl}
$\multiplication$ & = &
\hspace{-1em}\multirow{6}{*}{
\begin{tabular}{lllll}
\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
&& $\Zero$ & $\Rightarrow$ & $\lambda f.~\Zero$ \\
&& $\Suc{m_2}$ &$\Rightarrow$& $\lambda f.$ \\
&&
\multicolumn{3}{l}{
\begin{tabular}{lll}
& \textsf{let} & $x~=~\Yab~m_2~n~f$ \\
& \textsf{in} & $\mathsf{add}~x~n$
\end{tabular}
}
\end{tabular}
}
\\
&&\\
&&\\
&&\\
&&\\
&&\\
&&\\
$\mathsf{times}~M~N$ &=& $\Yab~M~N~\multiplication$
\end{tabular}
\end{center}

\section{Evaluation}

We computed some random numbers woo

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.48\textwidth]{rslc-rdr.pdf}
	\end{center}
	\caption{We show that RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
	\label{fig:rdr}
\end{figure}

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.40\textwidth]{rslc-d20.pdf}
	\end{center}
	\caption{RSLC can also be used to implement popular role-playing games.}
	\label{fig:rdr}
\end{figure}

>>>>

Unfortunately, I wasn't even able to find an RSLC term that satisfied that identity. I had to also alter the way recursive functions call themselves -- by referring to the combinator itself, passing it the recursive arguments, and {\em then} the substituted version of themselves. Since that makes no sense in words, just have a look:
\begin{eqnarray*}
\Ya &=& \lambda ag.~gag \\
\Yab &=& \lambda abg.~gabg \\
\Yabc &=& \lambda abcg.~gabcg \\
&\dots&
\end{eqnarray*}
Here's a sort of handwavy syntactic transformation example for $\mathsf{fix}$ constructs that supports one, two, etc arguments:
\begin{eqnarray*}
\mathsf{fix}_1(f,a.e) &=& \lambda a. \Ya~a~(\lambda af. [^{(\Ya~x~f)}/_{f(x)}]e) \\
\mathsf{fix}_2(f,a,b.e) &=& \lambda ab. \Ya~a~b~(\lambda abf. [^{(\Ya~x~y~f)}/_{f(x)(y)}]e) \\
&\dots&
\end{eqnarray*}

@@

\newcommand{\addition}{\maltese}
\begin{center}
\begin{tabular}{rcl}
$\addition$ & = &
\hspace{-1em}\multirow{4}{*}{
\begin{tabular}{lllll}
\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
&& $\Zero$ & $\Rightarrow$ & $\lambda f.~n$ \\
&& $\Suc{m'}$ &$\Rightarrow$& $\lambda f.~\Suc{\Yab~m'~n~f}$
\end{tabular}
}
\\
&&\\
&&\\
&&\\
&&\\
$\mathsf{add}~M~N$ &=& $\Yab~M~N~\addition$
\end{tabular}
\end{center}

@@

\newcommand{\multiplication}{\bigstar}
\newcommand{\multx}{\hat{x}}
\newcommand{\multm}{\hat{m}}
\newcommand{\multn}{\hat{n}}
\newcommand{\multf}{\hat{f}}
\newcommand{\multmm}{\hat{m}'}
\begin{center}
\begin{tabular}{rcl}
$\multiplication$ & = &
\hspace{-1em}\multirow{6}{*}{
\begin{tabular}{lllll}
\multicolumn{5}{l}{$\lambda \multm. \lambda \multn.$} \\
& \multicolumn{4}{l}{$\mathsf{case}~\multm~\mathsf{of}$} \\
&& $\Zero$ & $\Rightarrow$ & $\lambda \multf.~\Zero$ \\
&& $\Suc{\multmm}$ &$\Rightarrow$& $\lambda \multf.$ \\
&&
\multicolumn{3}{l}{
\begin{tabular}{lll}
& \textsf{let} & $\multx~=~\Yab~\multmm~\multn~\multf$ \\
& \textsf{in} & $\mathsf{add}~\multx~\multn$
\end{tabular}
}
\end{tabular}
}
\\
&&\\
&&\\
&&\\
&&\\
&&\\
&&\\
$\mathsf{times}~M~N$ &=& $\Yab~M~N~\multiplication$
\end{tabular}
\end{center}
Note that since $\mathsf{times}$ refers to $\mathsf{add}$, $\mathsf{times}$'s variable names must have different names. I put hats on them.

Here's the factorial function:
\newcommand{\factorial}{\textbf{!}}
\newcommand{\factx}{\mathring{x}}
\newcommand{\factn}{\mathring{n}}
\newcommand{\factf}{\mathring{f}}
\newcommand{\factnn}{\mathring{n}'}
\begin{center}
\begin{tabular}{rcl}
$\factorial$ & = &
\hspace{-1em}\multirow{6}{*}{
\begin{tabular}{lllll}
\multicolumn{5}{l}{$\lambda \factn.$} \\
& \multicolumn{4}{l}{$\mathsf{case}~\factn~\mathsf{of}$} \\
&& $\Zero$ & $\Rightarrow$ & $\lambda \factf.~\Suc{\Zero}$ \\
&& $\Suc{\factnn}$ &$\Rightarrow$& $\lambda \factf.$ \\
&&
\multicolumn{3}{l}{
\begin{tabular}{lll}
& \textsf{let} & $\factx~=~\Ya~\factnn~\factf$ \\
& \textsf{in} & $\mathsf{times}~\factn~\factx$
\end{tabular}
}
\end{tabular}
}
\\
&&\\
&&\\
&&\\
&&\\
&&\\
&&\\
$\mathsf{fact}~N$ &=& $\Ya~N~\factorial$
\end{tabular}
\end{center}

\newcommand{\fst}{\mathsf{fst}}
\newcommand{\snd}{\mathsf{snd}}
To compute fibonacci numbers, I encoded pairs:
\begin{eqnarray*}
(e_1, e_2) &=& \lambda b.~b~e_1~e_2 \\
\fst~e &=& e (\lambda x.\lambda y. x) \\
\snd~e &=& e (\lambda x.\lambda y. y) \\
\end{eqnarray*}
Hence:
\newcommand{\fib}{\Phi}
\newcommand{\fibxa}{\tilde{x}_1}
\newcommand{\fibxb}{\tilde{x}_2}
\newcommand{\fibn}{\tilde{n}}
\newcommand{\fibf}{\tilde{f}}
\newcommand{\fibp}{\tilde{p}}
\newcommand{\fibnn}{\tilde{n}'}
\begin{center}
\begin{tabular}{rcl}
$\fib$ & = &
\hspace{-1em}\multirow{6}{*}{
\begin{tabular}{lllll}
\multicolumn{5}{l}{$\lambda \fibn.$} \\
& \multicolumn{4}{l}{$\mathsf{case}~\fibn~\mathsf{of}$} \\
&& $\Zero$ & $\Rightarrow$ & $\lambda \fibf.~(\Zero,~\Suc{\Zero})$ \\
&& $\Suc{\fibnn}$ &$\Rightarrow$& $\lambda \fibf.$ \\
&&
\multicolumn{3}{l}{
\begin{tabular}{lll}
& \textsf{let} & $\fibp~=~\Ya~\fibnn~\fibf$ \\
& & $\fibxa~=~\fst~\fibp$ \\
& & $\fibxb~=~\snd~\fibp$ \\
& \textsf{in} & $(\fibxb,~\mathsf{add}~\fibxa~\fibxb)$
\end{tabular}
}
\end{tabular}
}
\\
&&\\
&&\\
&&\\
&&\\
&&\\
&&\\
&&\\
&&\\
$\mathsf{fib}~N$ &=& $\Ya~N~\fib$
\end{tabular}
\end{center}

\section{Applications}

I computed some random numbers woo

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.48\textwidth]{rslc-rdr.pdf}
	\end{center}
	\caption{RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
	\label{fig:rdr}
\end{figure}

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.40\textwidth]{rslc-d20.pdf}
	\end{center}
	\caption{RSLC can also be used to implement popular role-playing games.}
	\label{fig:d20}
\end{figure}
